---
title: "R Notebook"
output: html_notebook
---

We can use the `read.delim()` function to import our `txt` file with Robert Frost's "The Road Not Taken":

```{r}
frostRoads <- read.delim("frostRoads.txt", header = FALSE, stringsAsFactors = FALSE)
```

We need to take some extra steps for cleaning the poem. First, it is imported as a data frame with a single variable named `V1`. We want to change that to `text` to match the code from [*Tidy Text Mining with `R`*](http://tidytextmining.com). We also want to cut out the line of code meant for managing chapters.

```{r}
frostRoads %>%
  rename(text = V1) %>%
  mutate(linenumber = row_number()) %>%
  ungroup() %>%
  unnest_tokens(word, text) -> tidyRoads
```

Our second block of code uses 2 lines instead of 80, since poems are quiet a bit shorter than novels. We also take out the reference to the `book` variable since that is not a relevant piece of data here.

```{r}
tidyRoads %>%
  inner_join(get_sentiments("bing")) %>%
  count(index = linenumber %/% 2, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative) -> sentimentRoads
```

Our plot goes unchanged except for updating the object meant to be plotted.

```{r}
ggplot(sentimentRoads, aes(x = index, y = sentiment)) +
  geom_col(fill = "#F8766D") +
  labs(
    title = "The Road Not Taken by Robert Frost",
    subtitle = "Sentiment Analysis")
```

Now you try to import and analyze another Frost poem, "Christmas Trees".

```{r}
frostXmas <- read.delim("frostXmas.txt", header = FALSE)

frostXmas %>%
  rename(text = V1) %>%
  mutate(text = as.character(text)) %>%
  mutate(linenumber = row_number()) %>%
  ungroup() %>%
  unnest_tokens(word, text) -> tidyXmas

tidyXmas %>%
  inner_join(get_sentiments("bing")) %>%
  count(index = linenumber %/% 2, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative) -> sentimentXmas

ggplot(sentimentXmas, aes(x = index, y = sentiment)) +
  geom_col(fill = "#F8766D") +
  labs(
    title = "Christmas Trees by Robert Frost",
    subtitle = "Sentiment Analysis")

```
